---
title: "pop-density-graphs"
author: "Meredith Shephard"
date: "10/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. import libraries

```{r}
library(tidyverse)
library(dplyr)
library(readxl)
library(broom)

```

2. import datasets

```{r}
pop <- read_csv("../../data/cleaned/USA_Counties_Data.csv")
jhu_cnty_death <- read_csv("../../data/raw/20211208_time_series_covid19_deaths_US.csv")
```

3. format and select population data columns

```{r}
pop$COUNTY_NAME <- toupper(pop$COUNTY_NAME)
pop$STATE_NAME <- toupper(pop$STATE_NAME)
pop <- pop %>%
  dplyr::select("COUNTY_NAME", "STATE_NAME", "POPULATION", "POP_SQMI")

```

Functions to create desired population density graphs:

4.0 - population density vs. cummulative covid-19 deaths for ALL US counties - no need to change

```{r}
all_us_death <- function(us_date1, us_date2){
  #rename columns
  names(jhu_cnty_death)[names(jhu_cnty_death)=="Admin2"] <- "COUNTY"
  names(jhu_cnty_death)[names(jhu_cnty_death)=="Province_State"] <- "STATE"
  jhu_cnty_death$COUNTY <- toupper(jhu_cnty_death$COUNTY)
  jhu_cnty_death$STATE <- toupper(jhu_cnty_death$STATE)
  
  us_death_long <- gather(jhu_cnty_death, DATE, CUMDEATHS, us_date1:us_date2, factor_key=TRUE)
  
  #death_long$DATE <- as.Date(death_long$DATE, format="%m/%d/%y")
  
  us_death_long <- us_death_long %>%
    select("COUNTY", "STATE", "DATE", "CUMDEATHS")%>%
    filter(DATE==us_date2)
  
  us_death_wpop <- us_death_long %>%
    left_join(pop, by=c("COUNTY"="COUNTY_NAME", "STATE"="STATE_NAME"))
  
  us_death_wpop <-us_death_wpop %>%
  mutate(NormCumDeaths = round(CUMDEATHS/(POPULATION/50000)))
  
  us_graph <- ggplot(data = us_death_wpop, aes(x=log(POP_SQMI), y=NormCumDeaths)) +
  geom_point()+
  geom_smooth(method=lm)+
  labs(x="Log of County Population Density", y="County Cummulative Deaths (per 50,000 ppl)", title = "Normalized Cum Deaths vs Pop. Density for all US counties") 
}
```

**4.1 parameters for 4.0 - choose date range 

```{r}
us_date1 = "3/1/20"
us_date2 = "4/12/21"

us_graph <- all_us_death(us_date1, us_date2)

us_graph
```



5.0 - population density vs cumulative covid-19 death regression graphs for each individual state 

```{r}
death <- function(state, date1, date2){
  #rename columns
  names(jhu_cnty_death)[names(jhu_cnty_death)=="Admin2"] <- "COUNTY"
  names(jhu_cnty_death)[names(jhu_cnty_death)=="Province_State"] <- "STATE"
  jhu_cnty_death$COUNTY <- toupper(jhu_cnty_death$COUNTY)
  jhu_cnty_death$STATE <- toupper(jhu_cnty_death$STATE)
  pop$COUNTY_NAME <- toupper(pop$COUNTY_NAME)
  pop$STATE_NAME <- toupper(pop$STATE_NAME)
  
  death_long <- gather(jhu_cnty_death, DATE, CUMDEATHS, date1:date2, factor_key=TRUE)
  
  #death_long$DATE <- as.Date(death_long$DATE, format="%m/%d/%y")
  

  
death_long_date1 <- death_long %>%
    dplyr::select("COUNTY", "STATE", "DATE", "CUMDEATHS")%>%
    filter(DATE==date1 & STATE==state)

death_long_date2 <- death_long%>%
  dplyr::select("COUNTY", "STATE", "DATE", "CUMDEATHS")%>%
  filter(DATE==date2 & STATE==state)

names(death_long_date2)[names(death_long_date2)=="CUMDEATHS"] <- "CUMDEATHS2"

new_death_long <- death_long_date1%>%
  left_join(death_long_date2, by=c("COUNTY", "STATE"))%>%
  mutate(NewCumDeaths = round(CUMDEATHS2 - CUMDEATHS))
   
  
  death_wpop <- new_death_long %>%
    left_join(pop, by=c("COUNTY"="COUNTY_NAME", "STATE"="STATE_NAME"))
  
  death_wpop <-death_wpop %>%
    mutate(NormCumDeaths = round(NewCumDeaths/(POPULATION/50000)),
           log_pop_density = log(POP_SQMI))%>%
    na.omit()
  
  
  
  regression <- lm(NormCumDeaths~log(POP_SQMI), data=death_wpop)
  
  state_graph <- ggplot(data = death_wpop, aes(x=log(POP_SQMI), y=NormCumDeaths)) +
  geom_point()+
  geom_smooth(method=lm)+
 labs(x="Log of County Population Density", y="County Cummulative Deaths (per 50,000 ppl)",  
       subtitle=paste(" Slope =",signif(regression$coef[[2]], 5)))+
    ggtitle(paste("Normalized Cumulative Deaths vs Population Density"))
  summary(regression)
}
```


**5.1 - parameters for 5.0- choose date range: 

```{r}
state = "WEST VIRGINIA"
date1 = "6/7/20"
date2 = "3/31/21"

state_graph <- death(state, date1, date2)


state_graph

write_csv(state_graph, paste(state, "_stats_june7.csv"))

#ggsave(paste(state,"_death_by_pop_figure.png"), state_graph, height=5, width=7)
```


6.0 - Create data frame for pop density stats

```{r}
death <- function(date1, date2){
  #rename columns
  names(jhu_cnty_death)[names(jhu_cnty_death)=="Admin2"] <- "COUNTY"
  names(jhu_cnty_death)[names(jhu_cnty_death)=="Province_State"] <- "STATE"
  jhu_cnty_death$COUNTY <- toupper(jhu_cnty_death$COUNTY)
  jhu_cnty_death$STATE <- toupper(jhu_cnty_death$STATE)
  pop$COUNTY_NAME <- toupper(pop$COUNTY_NAME)
  pop$STATE_NAME <- toupper(pop$STATE_NAME)
  
  death_long <- gather(jhu_cnty_death, DATE, CUMDEATHS, date1:date2, factor_key=TRUE)
  
  #death_long$DATE <- as.Date(death_long$DATE, format="%m/%d/%y")
  

  
death_long_date1 <- death_long %>%
    dplyr::select("COUNTY", "STATE", "DATE", "CUMDEATHS")%>%
    filter(DATE==date1)

death_long_date2 <- death_long%>%
  dplyr::select("COUNTY", "STATE", "DATE", "CUMDEATHS")%>%
  filter(DATE==date2)

names(death_long_date2)[names(death_long_date2)=="CUMDEATHS"] <- "CUMDEATHS2"

new_death_long <- death_long_date1%>%
  left_join(death_long_date2, by=c("COUNTY", "STATE"))%>%
  mutate(NewCumDeaths = round(CUMDEATHS2 - CUMDEATHS))
   
  
  death_wpop <- new_death_long %>%
    left_join(pop, by=c("COUNTY"="COUNTY_NAME", "STATE"="STATE_NAME"))
  
  death_wpop <-death_wpop %>%
    mutate(NormCumDeaths = round(NewCumDeaths/(POPULATION/50000)))%>%
    na.omit()%>%
    filter(STATE != "ALASKA" & STATE != "DISTRICT OF COLUMBIA")
  
  regression <- death_wpop %>%
    group_by(STATE)%>%
    do(tidy(lm(NormCumDeaths~log(POP_SQMI), data = .)))%>%
    slice(2)
    
  
}
```

**6.1 - parameters - choose date range for analysis and change output name:

```{r}
date1 = "5/1/20"
date2 = "3/31/21"

regression_stats <- death(date1, date2)

regression_stats

write_csv(regression_stats, "regression_stats_may1.csv")
```


7. Create Dot plot for pop density slope 

```{r}
pop_slope <- read_csv("../../../data/inputs/st-popdensity-slope-data.csv")
library(ggthemes)
col_theme <- c("Positive"="#009E73","No Relationship"="#0072B2","Negative"="#D55E00")

pop_slope_dotplot <- pop_slope%>%
  drop_na(POP_SLOPE)%>%
  ggplot(aes(POP_SLOPE))+
  geom_dotplot(aes(y=log(POP_SQMI), fill=POP_SLOPE), binaxis = 'y', stackdir='center')+
  guides(fill=FALSE)+
  labs(y="Log of Population Density", x = "Relationship between Pop Density and Cummulative COVID Deaths", title="Log of Population Density vs Pop Density Relationship")+
  scale_fill_manual(values=col_theme)

pop_slope_dotplot

ggsave("pop-slope-dotplot.png", pop_slope_dotplot, height=7, width=9)
```



