1. import libraries and and dataset - no need to change
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(readxl)
library(ggplot2)
library(dplyr)
library(plotly)
library(ggseas)
library(chron)
```

read and select data
```{r, warning=FALSE,message=FALSE}
data <- read_csv("../../data/cleaned/NPI_cases_04162021.csv") %>% 
  select(c("STATE", "DATE", "NEWCONFIRMED", "EVENT_CATG_S",
           "SCORE", "SHORT_DESCRIPT"))
state_pop_dt <- read_csv("../../data/cleaned/USA_States_Data.csv") %>% 
  select(c("STATE_NAME", "POPULATION"))
```


2. preprocess data - join population and covid data
```{r}
# change state population data state_name to all caps
state_pop_dt$STATE_NAME <- toupper(state_pop_dt$STATE_NAME)

# change colum names
colnames(data) <- c("State", "Date", "New_Confirmed", "Event_Catg_S","Score", "Short_Description")
colnames(state_pop_dt) <- c("State", "Population")

# join covid and populaton data
data <- inner_join(data, state_pop_dt, by=c("State")) 
data$Date <- as.Date(data$Date, format="%m/%d/%Y")
```

rescale data into 500 normalized
```{r}
data$`New Cases/500,000 ppl`<-
  round(data$New_Confirmed/(data$Population/500000))

```


preprocess data - event categories
```{r}
# add score line
data$dummy <- -10

# change category events name
# group event categories into easing/lifting and imposing
create_melr_catg <- function(x){
  new_col = c()
  check = substr(x, nchar(x), nchar(x))
  if (is.na(check)){
    new_col <- append(new_col, x)
  } else if(check %in% c("M","A","I", "R")){
    new_col <- append(new_col, "Restricting")
  } else if (check == "R"){
    new_col <- append(new_col, "Restricting")
  } else if (check == "E"){
    new_col <- append(new_col, "Easing")
  } else if (check == "L"){
    new_col <- append(new_col, "Lifting")
  } 
}
data$event_MELR <- sapply(data$Event_Catg_S,
                                    FUN=create_melr_catg)
# category num
create_event_catg <- function(x){
  new_col = c()
  if (is.na(x)){
    new_col <- append(new_col, x)
  } 
  else if(grepl("15", x)){
    new_col <- append(new_col, "Face\nCovering")
  } else if(grepl("2", x)){
    new_col <- append(new_col, "Business")
  } else if(grepl("3", x)){
    new_col <- append(new_col, "Indoor Gatherings")
  } else if(grepl("4", x)){
    new_col <- append(new_col, "Restaurant/Bar")
  } else{
    new_col <- append(new_col, "Stay at\nHome")
  }
}

data$big_cat <- sapply(data$Event_Catg_S, FUN=create_event_catg)
data$Event_ <- paste(data$big_cat, "\n", data$event_MELR, sep="")
data$Event_[data$Event_=="NA\nNA"] <- NA



```
```{r}
# remove rows if date is NA
data <- data[!is.na(data$Date),]
# find 9/7 for labor day
data[data$Date=="2020-09-07",]$Event_ <- "Labor Day"
# find 12/11 for Pfizer vaccine
data[data$Date=="2020-12-11",]$Event_ <- 
  "Pfizer Vaccine\nApproved for\nUSA Distribution"
data[data$Date=="2020-12-18",]$Event_ <-
  "Moderna Vaccine\nApproved for\nUSA Distribution"
data$Event <- gsub("\r?\n|\r", " ", data$Event_) # show up in tooltip
```

```{r}
# text color bins
bins <- c(-Inf,0,0.55,1.20,1.70, 2.15,2.60,3.10,3.60,4.3, 5.00)
data$color_scale <- cut(data$Score, breaks=bins)
color_v <- c("#ffffff","#fdfec0", "#fedc9e","#fe9969","#f9765d", "#ec5761",
             "#b7377a","#5f1888","#231252","#000004") 

data$Event_ <- as.factor(data$Event_)
color_event_v <- c("Business\nEasing"="#44aa99",
                   "Business\nLifting"="#44aa99",
                   "Business\nRestricting"="#aa4499",
                   "Business\nRestricting"="#aa4499",
                   "Face\nCovering\nRestricting"="#aa4499",
                   "Face\nCovering\nLifting"="#44aa99",
                   "Indoor Gatherings\nEasing"="#44aa99",
                   "Indoor Gatherings\nLifting"="#44aa99",
                   "Indoor Gatherings\nRestricting"="#aa4499",
                  "Indoor Gatherings\nRestricting"="#ad1711",
                  "Restaurant/Bar\nEasing"="#44aa99",
                  "Restaurant/Bar\nLifting"="#44aa99",
                  "Restaurant/Bar\nRestricting"="#aa4499",
                  "Restaurant/Bar\nRestricting"="#a60903",
                  "Stay at\nHome\nLifting"="#44aa99",
                  "Stay at\nHome\nRestricting"="#aa4499",
                  "Labor Day"="black",
                  "Moderna Vaccine\nApproved for\nUSA Distribution"="black",
                  "Pfizer Vaccine\nApproved for\nUSA Distribution"="black")


names(color_v) = levels(data$color_scale)


text_y_pos <- function(state_event, days_diff, space_decs,
                       label_position_offset){
  y_pos_vec = c(max(
    state_event[,"New Cases/500,000 ppl"])+label_position_offset)
  for (i in 2:nrow(state_event)){
    diff = state_event[i,"Date"] - state_event[i-1,"Date"]
    if (as.integer(diff)<days_diff){
      tmp = y_pos_vec[i-1]-space_decs
       y_pos_vec = append(y_pos_vec, tmp)
    } else {
      y_pos_vec = append(y_pos_vec,
                         max(state_event[,"New Cases/500,000 ppl"])+
                           label_position_offset)
    }
  }
  return(y_pos_vec)
}

```


3. Create an interactive version for the time trend - no need to change
```{r message=FALSE, warning=FALSE}

plot_time_interact <- function(state, window_num, days_diff, space_desc,
                   label_position_offset, y_axis_max,
                   no_score_legend_lower, no_score_legend_upper){
  state_data <- data[data$State==state,]
  state_event <- state_data[!is.na(state_data$Event_),]
  state_event_date <- state_event$Date
  state_event$txt_y <- text_y_pos(state_event, days_diff, space_decs,
                       label_position_offset)

  p <- ggplot(state_data, aes(Date, `New Cases/500,000 ppl`,
                              label1=Date)) +
    geom_point(aes(label2=`New Cases/500,000 ppl`),size=2, alpha=1) + 
    stat_rollapplyr(width = window_num, align = "right", size = 2,
                    col="gray", alpha=1)+
    geom_point(mapping=aes(x=Date, y=dummy, fill=color_scale,
                           label3=Score), shape=22, size=2,
               color="transparent")+
    scale_fill_manual(values=color_v, name="")+
    scale_color_manual(values=color_event_v, name="")+
    geom_vline(data=state_event,
               aes(xintercept=as.numeric(Date), color=Event_), 
               linetype="dotted")+
    geom_text(data=state_event, aes(x=Date, y=txt_y, color=Event_,
                                    label=paste("<b>",
                                                as.character(Event_),
                                                "</b>"),
                                    label4=Event,
                                    label5=Short_Description),
              size=2.75)+
    ylim(-15, y_axis_max)+
    theme(legend.title=element_blank())+
    theme_bw()+

    labs(x = "Months", y = "New Daily Cases Normalized")+
    scale_x_date(date_breaks = "months" , date_labels = "%b")+
    ggtitle(paste(state, ",",window_num, "Days Rolling Average")) 
    
  interactive_p <- ggplotly(p,
                     tooltip=c("label1","label2","label3","label4",
                               "label5")) %>% 
    layout(legend=list(title=list(text='<b> Orders</b>')))
  
  interactive_p<- style(interactive_p, showlegend = FALSE,
                 traces=c(no_score_legend_lower:no_score_legend_upper))
  
  for (i in 1:length(interactive_p$x$data)){
    if (!is.null(interactive_p$x$data[[i]]$name)){
        interactive_p$x$data[[i]]$name = 
          gsub("\\(","",str_split(interactive_p$x$data[[i]]$name,",")[[1]][1])
    }
}
  interactive_p

}

```



***** 4. run interactive time series plot with specified state and window
change variables for different states, moving average window, and aesthetic
```{r, message=FALSE, warning=FALSE}
# identify which individual state
state = "WYOMING"
# identify the window(days) for moving averages
window_num = 7

# the next four variables are for aesthetic only 
# changing these variables can lesson the problems of text overlapping

# if days_diff is too small, all the neighboring intervention texts will overlap
days_diff = 26
# identify how much we want to separate this intervention text from last
# intervention text
space_decs = 75
# y position of highest text label
label_position_offset = 260
# y axis max height
y_axis_max = 900
no_score_legend_lower = 1 #no need to change
no_score_legend_upper = 8 #start with 7 or smaller, if score legend included, gradually increase

# function to plog interactive time series - no need to change

p <- plot_time_interact(state, window_num, days_diff, space_desc,
                   label_position_offset, y_axis_max,
                   no_score_legend_lower, no_score_legend_upper)
htmlwidgets::saveWidget(as_widget(p),
                        paste(tolower(state),"_interactive.html",sep=""))
```




